\section{Acknowledgements}

As I was writing my thesis I spent a lot of time
    thinking about all the people who helped me along the way
    and feeling an immense amount of gratitude.
During this process it seems silly that we write such short preface to such a long
    document when the people, their time, energy, thoughtfulness, friendship,
    and roles seem so much larger than the things that ended up in this
    document.
I will inevitably forget to thank some people,
    if you feel that you helped me you probably did; here it goes.

First of all thanks to my family, Kevin, Adrienne, and Ian who have always encouraged
    me to be my own thinker and let me do it differently even if they don't always fully
    understand my intensity and insanity.
To my undergraduate mentors without whom I would probably be an unemployed
    jazz musician somewhere.
Murat Karaoraman who believed in me from day one even though
    I didn't have much to show yet.
If you hadn't bet on my potential
    I wouldn't be a computer scientist today.
To Jay Freeman, who showed me what a true
    hacker is.
To Ben Hardekopf and Tim Sherwood who taught me how to do research
    and seek out interesting questions with a PL spin.
To my UCSB labmates who helped me grow into a researcher and
    provided inspiration, wisdom, and friendship.
To Joseph who has remained a close friend after failing to
    convince me to go to the hospital during a deadline push,
    in the end I only lost an Appendix.
To the UCSB crew, Pete, Tristan, Andrew, Chang, and more who helped me grow
    as a hacker, and not settle for less.
To Daniel for staying up at all hours of the night talking about research,
    startups, or whatever stupid argument we had that week.

To UW PLSE, an amazing second home and group of people to
    go through graduate school.
The lab crew, and our times in the PLSE lab playing loud
    music, pranks, games, and banter.
Everyone was a friend and important part of my graduate
    school experience but specifically the people
    who were my closest friends (and sometimes roommates)
    during this time Max Willsey, Chandrakana Nandi,
    Sam Elliot, Luke Nelson, Shumo Chu, Chenglong Wang, and Luis Vega.
Most importantly to all the amazing people I got to work
    with as collaborators across all my projects.
Especially the Relay crew of Steven Lyubomirsky,
    Josh Pollock, Logan Weber, Marisa Kirisame, and Altan Haan.

To all my partners and friends during this time, who tolerated
    my obsession with work and shipping just one more 4k PR.
To Pavel Panchekha who told me "doing verification sucks",
    I eventually got the message a few years later.
To my co-founders at OctoML who brought me along for the ride
    even though I wasn't quite ready yet.
Thierry for grinding through the end of PhD together,
    fighting fires and ``growing up'' together.
Tianqi for diverting my life into an amazingly fun and
    different direction thanks to a random email.
Jason, for reminding me why betting on people is worth doing.
Luis for having a grand vision and pushing us to dream
    bigger.
Thanks to Amanda Robles for having our backs both in CSE
    and OctoML, you helped me numerous times even during
    the writing of this last section of my thesis.
To all my co-workers at OctoML its been amazing to build
    a company with you.
I appreciate the opportunity to learn and grow with such
    great people, and appreciate you taking a bet on us
    co-founders.

To Leo De Moura who was an amazing mentor and friend for the start of
    my PhD and taught me things about research, life, and writing lots of code.
To Zachary L. Tatlock my devious advisor who was always there
    to pull the chair out from under me both literally and figuratively
    while laughing about it.
I learned how to run a group, mentor, invest in people, attempt
    balance in a busy life, and do big things.
To Xi Wang who was my co-advisor for the early part of my PhD
    and always someone who had wisdom and advice for me.
I learned the importance of not being dogmatic in your approach or tools,
    and giving up early.
Finally to Eunice who has put up with me as I have struggled
    to have a social life, relationship, found and run a startup,
    while finishing this little thing called a PhD on the side.
Thanks for being my late night editor, sounding board, and
    partner in crime.
You were always the taste tester for my real job
    of cooking way too often in the past few years.
This document should really be titled the PhD Pigout, as I probably cooked
    and ate more then I did research.
Finally the most important star of the show that last 6 years: coffee.
This PhD wouldn't of been possible without the copious amount of coffee,
    and hours programming at coffee shops.
I think one place in particular may of been the location of nearly 100,000
    lines of Coq proofs early in my PhD.

I will always remember everyone named and unnamed as part of this awesome journey,
    thanks again to all for being in my life and helping me achieve my goals.
